Automatic Bicycle Balance Assistance Reduces Probability of Falling at Low Speeds When Subjected to Handlebar Perturbations\\Version 2.dev

### 2024-09-23 main.tex
Use builtin appendix feature and reference the appendix.

\subsection{Balance Assist Control}
%
We use a forward speed \(v\) gain scheduled proportional roll rate feedback
controller to stabilize the bicycle. In the speed range tested, the commanded
steer torque \(T^\textrm{m}_\delta\) from the steer motor follows the control
law
%
\begin{align}
  T^\textrm{m}_\delta = -k_{\dot{\phi}}\dot{\phi} = g(v_{stable} - v)\dot{\phi}
  \label{eq:implemented-controller}
\end{align}
%
where \(v_{\textrm{stable}} = 4.7~\si{\meter\per\second}\) is approximately the
average stable speed predicted from the open loop bicycle rigid-rider dynamics.
We use \(g=8\) and \(g=10\) as gain values during the experiments. Scaling the
proportional feedback gain linearly with respect to speed stabilizes the
normally unstable weave mode of the bicycle down to
2.9~\si{\kilo\meter\per\hour} for the riderless bicycle and
4.6~\si{\kilo\meter\per\hour} for the ridden bicycle  as shown in
Figure~\ref{fig:eigenvalues}.
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{figures/balance-assist-eig-vs-speeds.png}
  \caption{Uncontrolled (upper row) and controlled (lower row) root locus of
    eigenvalue components (real: solid, imaginary: dashed) plotted versus speed
    for the bicycle without the inertial effect of a rigid rider (left) and
    with a rigid rider (right). Vertical dotted and dotted-dashed lines
    indicate the two speeds we perform experiments at. Grey shaded region is
    the linear stable speed range. Geometric and inertial parameters for these
    plots were estimated with the methods presented in \cite{Moore2012} and
    software packages BicycleParameters 1.1.1~\citep{Moore2024} and Yeadon
    1.5.0~\citep{Dembia2015}.}
  \label{fig:eigenvalues}
\end{figure}

@@

\section*{Appendix}
%
Physical parameters and estimated uncertainty (standard deviation) for the
balance-assist bicycle without and with the rider. These are derived from the
measurement approach in \citet{Moore2012}. The rider inertial parameters are
derived from the rider ``Jason'' available in BicycleParameters
1.1.1~\citep{Moore2024}. The variable names match the notation from the
benchmark bicycle model presented in \citet{Meijaard2007}.

\begin{tabular}{llrrrr}
  \toprule
  Variable & Unit & Without Rider & Uncertainty & With Rider & Uncertainty \\
  \midrule
  \(c\) & \si{\meter} & 0.04 & 0.01 & 0.04 & 0.01 \\
  \(g\) & \si{\meter\per\second\squared} & 9.80665 & NA & 9.80665 & NA \\
  \(I_{Bxx}\) & \si{\kilogram\meter\squared} & 1.12 & 0.02 & 19.7 & 0.3 \\
  \(I_{Bxz}\) & \si{\kilogram\meter\squared} & 0.05 & 0.03 & -3.5 & 0.2 \\
  \(I_{Byy}\) & \si{\kilogram\meter\squared} & 3.16 & 0.08 & 22.7 & 0.4 \\
  \(I_{Bzz}\) & \si{\kilogram\meter\squared} & 2.12 & 0.03 & 5.0 & 0.1 \\
  \(I_{Fxx}\) & \si{\kilogram\meter\squared} & 0.0995 & 0.0005 & 0.0995 & 0.0005 \\
  \(I_{Fyy}\) & \si{\kilogram\meter\squared} & 0.1902 & 0.0009 & 0.1902 & 0.0009 \\
  \(I_{Hxx}\) & \si{\kilogram\meter\squared} & 0.298 & 0.003 & 0.298 & 0.003 \\
  \(I_{Hxz}\) & \si{\kilogram\meter\squared} & -0.038 & 0.007 & -0.038 & 0.007 \\
  \(I_{Hyy}\) & \si{\kilogram\meter\squared} & 0.257 & 0.006 & 0.257 & 0.006 \\
  \(I_{Hzz}\) & \si{\kilogram\meter\squared} & 0.057 & 0.003 & 0.057 & 0.003 \\
  \(I_{Rxx}\) & \si{\kilogram\meter\squared} & 0.1023 & 0.0006 & 0.1023 & 0.0006 \\
  \(I_{Ryy}\) & \si{\kilogram\meter\squared}& 0.1887 & 0.0007 & 0.1887 & 0.0007 \\
  \(\lambda\) & \si{\radian} & 0.25 & 0.03 & 0.25 & 0.03 \\
  \(m_B\) & \si{\kilogram} & 22.50 & 0.01 & 106.00 & 0.01 \\
  \(m_F\) & \si{\kilogram} & 2.23 & 0.01 & 2.23 & 0.01 \\
  \(m_H\) & \si{\kilogram} & 4.30 & 0.01 & 4.30 & 0.01 \\
  \(m_R\) & \si{\kilogram} & 4.08 & 0.01 & 4.08 & 0.01 \\
  \(r_F\) & \si{\meter} & 0.3523 & 0.0001 & 0.3523 & 0.0001 \\
  \(r_R\) & \si{\meter} & 0.3489 & 0.0001 & 0.3489 & 0.0001 \\
  \(w\) & \si{\meter} & 1.113 & 0.008 & 1.113 & 0.008 \\
  \(x_B\) & \si{\meter} & 0.52 & 0.01 & 0.374 & 0.002 \\
  \(x_H\) & \si{\meter} & 0.92 & 0.02 & 0.92 & 0.02 \\
  \(z_B\) & \si{\meter} & -0.52 & 0.02 & -1.009 & 0.003 \\
  \(z_H\) & \si{\meter} & -0.860 & 0.006 & -0.860 & 0.006 \\
  \bottomrule
\end{tabular}

>>>>

\subsection{Balance Assist Control}
%
We use a forward speed \(v\) gain scheduled proportional roll rate feedback
controller to stabilize the bicycle. In the speed range tested, the commanded
steer torque \(T^\textrm{m}_\delta\) from the steer motor follows the control
law
%
\begin{align}
  T^\textrm{m}_\delta = -k_{\dot{\phi}}\dot{\phi} = g(v_{stable} - v)\dot{\phi}
  \label{eq:implemented-controller}
\end{align}
%
where \(v_{\textrm{stable}} = 4.7~\si{\meter\per\second}\) is approximately the
average stable speed predicted from the open loop bicycle rigid-rider dynamics.
We use \(g=8\) and \(g=10\) as gain values during the experiments. Scaling the
proportional feedback gain linearly with respect to speed stabilizes the
normally unstable weave mode of the bicycle down to
2.9~\si{\kilo\meter\per\hour} for the riderless bicycle and
4.6~\si{\kilo\meter\per\hour} for the ridden bicycle  as shown in
Figure~\ref{fig:eigenvalues}.
%
\begin{figure}
  \centering
  \includegraphics[width=160mm]{figures/balance-assist-eig-vs-speeds.png}
  \caption{Uncontrolled (upper row) and controlled (lower row) root locus of
    eigenvalue components (real: solid, imaginary: dashed) plotted versus speed
    for the bicycle without the inertial effect of a rigid rider (left) and
    with a rigid rider (right). Vertical dotted and dotted-dashed lines
    indicate the two speeds we perform experiments at. Grey shaded region is
    the linear stable speed range. Geometric and inertial parameters for these
    plots were estimated with the methods presented in \cite{Moore2012} and
    software packages BicycleParameters 1.1.1~\citep{Moore2024} and Yeadon
    1.5.0~\citep{Dembia2015} and are shown in
    Appendix~\ref{sec:physical-parameters}.}
  \label{fig:eigenvalues}
\end{figure}

@@

\appendix
%
\section{Physical Parameters}
\label{sec:physical-parameters}
%
Physical parameters and estimated uncertainties (standard deviation) for the
balance-assist bicycle with and without a rider. These are found using the
measurement approach for geometric and inertial parameters for a bicycle
explained in \citet{Moore2012}. The rider inertial parameters are derived from
the 83.5~\si{\kilogram} rider ``Jason'' available in BicycleParameters
1.1.1~\citep{Moore2024}. The variable names match the notation and definitions
from the benchmark bicycle model presented in
\citet{Meijaard2007}.\todo{Uncertainties are based on best guess for
uncertainty in the raw measurements, may be better to remove them as they
aren't so meaningful.}
%
\begin{table}[h]  % have to use "h" so that it stays within the appendix
  \centering
  \caption{Physical parameters of the balance-assist bicycle with and without
    the inertia effects of a rigid rider.}
  \begin{tabular}{llrrrr}
    \toprule
    Variable & Unit & Without Rider & Uncertainty & With Rider & Uncertainty \\
    \midrule
    \(c\) & \si{\meter} & 0.04 & 0.01 & 0.04 & 0.01 \\
    \(g\) & \si{\meter\per\second\squared} & 9.80665 & NA & 9.80665 & NA \\
    \(I_{Bxx}\) & \si{\kilogram\meter\squared} & 1.12 & 0.02 & 19.7 & 0.3 \\
    \(I_{Bxz}\) & \si{\kilogram\meter\squared} & 0.05 & 0.03 & -3.5 & 0.2 \\
    \(I_{Byy}\) & \si{\kilogram\meter\squared} & 3.16 & 0.08 & 22.7 & 0.4 \\
    \(I_{Bzz}\) & \si{\kilogram\meter\squared} & 2.12 & 0.03 & 5.0 & 0.1 \\
    \(I_{Fxx}\) & \si{\kilogram\meter\squared} & 0.0995 & 0.0005 & 0.0995 & 0.0005 \\
    \(I_{Fyy}\) & \si{\kilogram\meter\squared} & 0.1902 & 0.0009 & 0.1902 & 0.0009 \\
    \(I_{Hxx}\) & \si{\kilogram\meter\squared} & 0.298 & 0.003 & 0.298 & 0.003 \\
    \(I_{Hxz}\) & \si{\kilogram\meter\squared} & -0.038 & 0.007 & -0.038 & 0.007 \\
    \(I_{Hyy}\) & \si{\kilogram\meter\squared} & 0.257 & 0.006 & 0.257 & 0.006 \\
    \(I_{Hzz}\) & \si{\kilogram\meter\squared} & 0.057 & 0.003 & 0.057 & 0.003 \\
    \(I_{Rxx}\) & \si{\kilogram\meter\squared} & 0.1023 & 0.0006 & 0.1023 & 0.0006 \\
    \(I_{Ryy}\) & \si{\kilogram\meter\squared}& 0.1887 & 0.0007 & 0.1887 & 0.0007 \\
    \(\lambda\) & \si{\radian} & 0.25 & 0.03 & 0.25 & 0.03 \\
    \(m_B\) & \si{\kilogram} & 22.50 & 0.01 & 106.00 & 0.01 \\
    \(m_F\) & \si{\kilogram} & 2.23 & 0.01 & 2.23 & 0.01 \\
    \(m_H\) & \si{\kilogram} & 4.30 & 0.01 & 4.30 & 0.01 \\
    \(m_R\) & \si{\kilogram} & 4.08 & 0.01 & 4.08 & 0.01 \\
    \(r_F\) & \si{\meter} & 0.3523 & 0.0001 & 0.3523 & 0.0001 \\
    \(r_R\) & \si{\meter} & 0.3489 & 0.0001 & 0.3489 & 0.0001 \\
    \(w\) & \si{\meter} & 1.113 & 0.008 & 1.113 & 0.008 \\
    \(x_B\) & \si{\meter} & 0.52 & 0.01 & 0.374 & 0.002 \\
    \(x_H\) & \si{\meter} & 0.92 & 0.02 & 0.92 & 0.02 \\
    \(z_B\) & \si{\meter} & -0.52 & 0.02 & -1.009 & 0.003 \\
    \(z_H\) & \si{\meter} & -0.860 & 0.006 & -0.860 & 0.006 \\
    \bottomrule
  \end{tabular}
\end{table}
